\documentclass[a4paper,12pt]{article}

\usepackage[a4paper,hmargin={2.5cm,1.5cm},vmargin={4cm,3cm}]{geometry}

\usepackage[english]{babel}
\usepackage{eurosym}
%\usepackage[latin1]{inputenc}
\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{indentfirst}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{rotfloat}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{indentfirst}
 \usepackage{dsfont}
 \usepackage{textcomp}
 \usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{pifont}
\usepackage{graphicx}
% for table
\usepackage{booktabs,caption} %,fixltx2e
\usepackage[flushleft]{threeparttable}
%http://tex.stackexchange.com/questions/12676/add-notes-under-the-table
% for citation
\usepackage[round, sort]{natbib}
\bibliographystyle{plainnat}
% for appendix
\usepackage[page]{appendix}
% for footnote in table
\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}
% table note
\usepackage{tablefootnote}
%http://tex.stackexchange.com/questions/68605/table-with-foot-note

%http://tex.stackexchange.com/questions/95866/multiple-references-to-the-same-footnote-inside-a-table-environment
% call `scrextend' before `hyperref'!
\usepackage{scrextend}
%\usepackage{hyperref}



 \long\def\symbolfootnote[#1]#2{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}
% \renewcommand{\baselinestretch}{1.2}
%\makeatletter
%\renewcommand\theequation{\thesection.\arabic{equation}}
%\@addtoreset{equation}{section} \makeatother
\font\dsrom=dsrom10 scaled 1200
\def \ind{\textrm{\dsrom{1}}}
\pagestyle{plain}

\begin{document}

\title{%
%TCIMACRO{\TeXButton{Begin single space}{\begin{singlespace}}}%
%BeginExpansion
%\begin{singlespace}%
%EndExpansion
The costs and benefits of collective reputation: Who gains and who loses from generic promotion programs?
%TCIMACRO{\TeXButton{End Single Space}{\end{singlespace}}}%
%BeginExpansion
%\end{singlespace}%
%EndExpansion
}

\author{
G. Cornelis van Kooten\thanks{%
University of Victoria, email: vkooten@uvic.ca} \\
%EndAName
  \and Jun Duan\thanks{%
University of Victoria, email: jonduan@uvic.ca} \\
%EndAName
  \and Rachel Lynch \thanks{%
University of Victoria, email: lrachel@uvic.ca} \\
%EndAName
} %\\
%EndAName


\maketitle

\begin{abstract}
%TCIMACRO{\TeXButton{Begin single space}{\begin{singlespace}}}%
%BeginExpansion
%\begin{singlespace}%
%EndExpansion
This paper explores the viability of relying on wind power to replace upwards of 60\% of electricity generation in Alberta that would be lost if coal-fired generation is phased out. Using hourly wind data from 17 locations across Alberta, we are able to simulate the potential wind power output available to the Alberta grid when modern, 3.5 MW-capacity wind turbines are spread across the province. Using wind regimes for the years 2006 through 2015, we find that available wind power is less than 60\% of installed capacity 98\% of the time, and below 30\% of capacity 74\% of the time. In addition, there is significant correlation between wind speeds at different locations, so it will always be necessary to rely on fossil fuel generation. Then, based on the results from a grid allocation model, we find that CO2 emissions can be reduced by about 30\%, but only through a combination of investment in wind energy and reliance on purchases of hydropower from British Columbia. Only if nuclear energy is permitted into the generation mix would Alberta be able to meet its CO2-emissions reduction target in the electricity sector. With nuclear power, emissions can be reduced by upwards of 85\%. 

\textbf{Key Words:} Electricity; renewable energy and climate change; wind power; intermittent energy; nuclear power.

\textbf{JEL Classification: }H41, L51, L94, Q42, Q48, Q54 
%TCIMACRO{\TeXButton{End Single Space}{\end{singlespace}}}%
%BeginExpansion
%\end{singlespace}%
%EndExpansion
\end{abstract}

\newpage

\baselineskip1.5em




\include{01-introduction}


%%
% Start the main matter (normal chapters)
%\mainmatter


\include{02-first-chapter}
\include{03-another-chapter}
\include{04-conclusion}




\newpage

%\backmatter

% for citation

\bibliography{references}

%\citet{Jones_2008} for Jones et al., (2008) and \citep{Jones_2008} for (Jones et al 2000) and no numbers or brackets

\include{appendix}


\end{document}
